---
title: "Source Augmentation Script ReadMe"
author: "Cynthia Yeh"
date: "April 12 2021"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Before getting started, the essential files are:

`r text_spec("1. In r folder", color = c("#79BAEC"), bold = T)`

augment.source 03032021_v3.R

source.functions.R

`r text_spec("1. In csv folder", color = c("#79BAEC"), bold = T)`

source_data

data_augmented_xl


Please make sure that there are no blanks in source.numbers column and the numbers are matched between Outcomes DB and Source DB. :)

note: The below example is from AMD source DB.


## DEFAULTS
### *Dependencies*
`r text_spec("Please install dplyr, purrr, tidyverse, lubridate before running the script.", background = "#E799A3", color ="white", bold = T)`

```{r install.pkg, message = FALSE}
#install.packages("dplyr")
#install.packages("purrr")
#install.packages("tidyverse")
#install.packages("lubridate")
library(dplyr)
library(purrr)
library(tidyverse)
library(lubridate)
```

### *Directories*
<br>Set up directories and make sure it is in the current database r folder

```{r}
getwd()
dirSource <- '../csv'
dirResults <- '../csv'
sourcedb <- 'source_data.csv'
augmentdb<-'data_augmented_xl.csv'
```

## DATA IMPORT
Import the data from csv folder and cleaning the Outcomes and Source DB.

```{r, include = FALSE}
OutcomesDB <- read.csv(file = paste(dirSource, augmentdb, sep = "/"), header = TRUE, fill = TRUE,
                       as.is = TRUE, sep = ",", na.strings = c(".","N/A"))
# exclude empty rows
OutcomesDB <- OutcomesDB %>%
  filter(!source.number == "")
SourceDB <- read.csv(file = paste(dirSource, sourcedb, sep = "/"), header = TRUE, fill = TRUE,
                     as.is = TRUE, sep = ",", na.strings = c(".","N/A"), strip.white = TRUE)
source('source.functions.R')
```

### *Check the original data dimensions*
```{r, results = "hold"}
dim(OutcomesDB)
dim(SourceDB)
```

## DATA PREP
```{r, include = FALSE}
SourceDB <- SourceDB %>%
  mutate(across(where(is.character), trimws))
```
### *Create SourceDB dataset*
```{r, include = FALSE}
### Dataset - raw source DB
datSource <- SourceDB
dim(datSource) 
```

### *Convert date columns to desired format using lubridate package*
```{r}
datSource$date.transform <- mdy(datSource$search.date)
```
```{r, echo = FALSE}
dates <- c("search.date", "date.transform")
for (k in dates) {
  datSource[,k] <- paste('="', as.character(datSource[,k]), '"', sep = '') 
}

datSource$search.date <- ifelse(datSource$date.transform == "=\"NA\"", 
                                as.character(datSource$search.date), 
                                datSource$date.transform)
```

### *Remove internal columns*

For example, the last 6 columns for the database are listed below: `r tail(names(datSource))`. Here, `r text_spec("entry", "qc", color = c("#79BAEC"), bold = T)` are internal columns that need to be removed.

`r text_spec("Please change colnames in internalCols to remove internal columns (ignore date.transform column).", background = "#E799A3", color ="white", bold = T)`
```{r}
internalCols <- c("entry", "qc")
```

```{r, include = FALSE}
datSource <- datSource %>%
  mutate(across(where(is.character), trimws),
         across(c(search, include, reason, e.copy, database), tolower)) %>%
  select(!all_of(internalCols) & !date.transform)
```

The updated columns are listed here: `r tail(names(datSource))`.
In most cases, `r text_spec("database", color = c("#79BAEC"), bold = T)` should be the last column in the source DB.

### *Subset Outcomes DB*
```{r}
OutcomesDB <- OutcomesDB %>%
  select(url:publication.type,
         study,
         registry.study.id,
         n.study,
         arm,
         control.arm,
         randomized.drug)
```

## DATA FORMAT CHECKS
In this section, we will check if there are any missing columns, unallowed non-numeric columns, and unify the columns format.

Example output as below (table 1):
```{r}
checkReqSourceCols(datSource)
```
`r text_spec("From the message above, please change the n.patients columns to n.study.", background = "#E799A3", color ="white", bold = T)`

```{r, echo = FALSE}
datSource <- datSource %>%
  rename(n.study = n.patients)
c1 <- names(datSource[1:12])
c2 <- c(names(datSource[13:23]), " ")
df<- cbind(c1, c2)
kable(df, format = "latex", booktabs = TRUE, caption = "Column Names") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
checkNumSourceCols(datSource)
```
`r text_spec("Uncomment the following 2 lines in the R-script to correct unallowed non-numerics.", background = "#E799A3", color ="white", bold = T, escape = FALSE)`

```{r}
#x <- c("source.number", "publication.year")
#datSource[x] <- sapply(datSource[x], as.numeric)
```

## NAMED RANGES
Check if the named ranges are matching (table 2). Please manually change them if there are unallowed names.

```{r, echo = FALSE, warning = FALSE}
df <- cbind(reasonType, searchType)

df[10:13,2] <- NA
df[1,1] <- "blanks"
opts <- options(knitr.kable.NA = "")
kable(df, format = "latex", booktabs = TRUE, caption = "Named Ranges") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

Example:
This chunk of codes will list out unallowed named ranges. Run it line by line.
```{r, results = "hide"}
unique(datSource$search[!datSource$search %in% searchType])
unique(datSource$include[!datSource$include %in% yesNoUnclear])
unique(datSource$reason[!datSource$reason %in% reasonType])
unique(datSource$e.copy[!datSource$e.copy %in% yesNo])
unique(datSource$database[!datSource$database %in% yesNo])
```

## URL HARMONIZATION
It makes the pubmed/clinical.trial urls consistent.

```{r, include = FALSE}
# Generate PubMed/clintrials URLs
datSource$url[datSource$search == "pubmed"] <- paste0("https://www.ncbi.nlm.nih.gov/pubmed/",
                                                      datSource$source.number[datSource$search == "pubmed"])
datSource$url[datSource$search == "clinicaltrials.gov"] <- paste0("https://clinicaltrials.gov/ct2/show/",
                                                                  datSource$registry.study.id[datSource$search == "clinicaltrials.gov"])
# Remove everything that comes after a comma in the URL
datSource$url <- sapply(strsplit(datSource$url, ","), "[", 1)
```

## REFERENCES CHECKS
Two messages to read in this section.

1st message: Whether there are references marked as database = yes in Source DB, but they are actually not found in Outcomes DB.
```{r, include = FALSE}
wDB.sNumber <- unique(OutcomesDB$source.number)
sDB.database.yes <- unique(datSource$source.number[datSource$database == "yes"])
sDB.database.no <- unique(datSource$source.number[datSource$database != "yes"])

# Check database = yes in source DB
wDB.no <- sDB.database.yes[!sDB.database.yes %in% wDB.sNumber]
```

2nd message: Check if all included references are curated or not.
```{r}
checkSourceYes(wDB.no)
```

```{r, include = FALSE}
# Check database == no in source DB
wDB.yes <- sDB.database.no[sDB.database.no %in% wDB.sNumber]
datSource$database[datSource$source.number %in% wDB.yes] <- "yes"

# Check the rest of database == no (& include == yes) in Source DB
sDB.include.yes <- datSource %>%
  filter(include == "yes", database != "yes") %>%
  select(source.number)
```

```{r}
checkSourceNo(sDB.include.yes)
```


## CREATE DATASETS
This section will reshape the dataset, check if we have `r text_spec("any references that have more than 1 study", color = c("#79BAEC"), bold = T)`, and will also combine the information for multiple studies in single references.
```{r, include = FALSE}
### Dataset - unique referenes that contains maximum arm for each study
datUniRef <- TrimData(OutcomesDB, source.number, study, arm)
datUniRef$arm <- gsub(".*\\+|[^0-9]", "", datUniRef$arm)
datUniRef <- datUniRef %>%
  group_by(source.number, study) %>%
  slice_max(arm)
```

Identify the source numbers for multiple studies references
```{r}
checkMultiStudy(datUniRef)
```

```{r, include = FALSE}
refs.multiple.study <- unique(datUniRef$source.number[duplicated(datUniRef$source.number)])

# Combine multiple studies information
datUniRef <- datUniRef %>%
  group_by(source.number) %>%
  mutate(across(c(study, registry.study.id, n.study, arm),
                function(x) paste(unique(x[!is.na(x) & x != ""]), collapse = " and "))) %>%
  filter(!duplicated(source.number))

### Dataset - unique control drugs dataset
datUniCtrl <- filter(OutcomesDB, !grepl("[A-z]", arm),            # exclude subarms
                     control.arm == "yes")                   # include control arms only
datUniCtrl <- TrimData(datUniCtrl, source.number, study, randomized.drug)

datUniCtrl <- datUniCtrl %>%
  group_by(source.number) %>%
  filter(!duplicated(randomized.drug)) %>%
  mutate(control.join = paste(randomized.drug, collapse = ", ")) %>%
  filter(!duplicated(source.number)) %>%
  select(source.number, control.join)

### Dataset - unique interventional drugs dataset
datUniTrt <- filter(OutcomesDB, !grepl("[A-z]", arm),             # exclude subarms
                    control.arm == "no")                     # include experimental arms only
datUniTrt <- TrimData(datUniTrt, source.number, study, randomized.drug)

datUniTrt <- datUniTrt %>%
  group_by(source.number) %>%
  filter(!duplicated(randomized.drug)) %>%
  mutate(treatment.join = paste(randomized.drug, collapse = ", ")) %>%
  filter(!duplicated(source.number)) %>%
  select(source.number, treatment.join)
```

After data wrangling, the first dimension of `r text_spec("datUniRef", color =c("#79BAEC"), bold = T)` and `r text_spec("datUniTrt", color =c("#79BAEC"), bold = T)` datasets should be the same.
```{r}
dim(datUniRef)
dim(datUniCtrl)
dim(datUniTrt)
```

## FILL IN MISSING INFORMATION
This part, we will overwrite Source database by gathering information from augmented Outcomes database. After this step, data be lined up between databases.
```{r, include = FALSE}
# Merge datUniRef/datUniCtrl/datUniTrt
cols.join <- c("url",
               "authors",
               "publication.year",
               "title",
               "journal",
               "volume",
               "pages",
               "study",
               "registry.study.id",
               "n.study",
               "arm")

datUniRef.join <- datUniRef %>%
  select(source.number, all_of(cols.join)) %>%
  rename(n.arm = arm) %>%
  rename_at(vars(-source.number), paste0, ".join") %>%
  list(datUniCtrl, datUniTrt) %>%
  reduce(left_join, by = "source.number")
```

Some databases, oncology database especially, there is no control arm due to single arm study design. Here, we will fill in control column of the Source DB as `r text_spec("no control", color =c("#79BAEC"), bold = T)`.
```{r}
datUniRef.join$control.join[is.na(datUniRef.join$control.join)] <- "no control"
```

```{r, include = FALSE}
### Check if it's in the augmented database, but not in the source database ###
miss.in.source <- wDB.sNumber[!wDB.sNumber %in% datSource$source.number]

# Add missing refs' source numbers to datUniRef.join for filling information later  
for (miss.info in miss.in.source) {
  datSource <- datSource %>%
    add_row(source.number = miss.info, database = "yes")
}


### All augmentated DB source.number should be in the source database now, proceed to fill in information ###
cols.join <- gsub(".join", "", names(datUniRef.join[, names(datUniRef.join) != "source.number"]))
cols.join

datSource <- datSource %>%
  left_join(datUniRef.join, by = "source.number")

# Fill in info from augmented outcome DB
for(names in cols.join) {
  datSource[[names]] <- ReplaceInfo(names)
}
```



## CHECK INFORMATION FOR REFERENCES MORE THAN 1 STUDY
A glimpse of combined information for multiple studies references.
```{r, include = FALSE}
check.multi.study <- datSource %>%
  filter(source.number %in% refs.multiple.study) %>%
  select(source.number,
         study, 
         registry.study.id,
         treatment,
         control,
         n.study,
         n.arm)
```
```{r, echo = FALSE}
view(check.multi.study)
```

## DATA FINAL CLEANING
Finally, we clean the data again and write the output to `r text_spec("source_augmented.csv", color =c("#79BAEC"), bold = T)` file.
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#Unify study name format (remove quotation marks)
datSource$study <- gsub('=|"', "", datSource$study)

datSource <- select(datSource, !ends_with("join"))
write.csv(datSource, paste(dirResults,'source_augmented.csv',sep = '/'), na = '', row.names = FALSE)
```


## MESSAGES
Just a second. we still have a few more friendly reminders here. In this section, we can check if there are any more filling needs to be done manually.
```{r, echo = FALSE}
ManualFill(miss.in.source)
MissingURLs(datSource)
```

The augmentation is complete. Thanks! :)
